Approximation as Differentiation

### 2025-06-09 main/first-order.tex
Now paring back LatGal.

\todo{Rework to account for new ordering.} Galois connections are pairs of monotone functions $f: Y \to X$ and
$g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse of $g$,
and $g$ the best approximation from above to an inverse of $f$. The category $\LatGal$ of bounded lattices and
Galois connections is semi-additive.

\begin{definition}[Galois connection]
Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
$f: Y \to X$ and $g: X \to Y$ satisfying $y \leq g(x) \iff f(y) \leq x$ for any $x \in X$ and $y \in Y$.
\end{definition}

\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
adjunction between poset categories, with monotone $f$ and $g$ regarded as functors; $f$ is usually referred
to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections compose
component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition.

>>>>

The category $\LatGal$ of bounded lattices and Galois connections is semi-additive.

\begin{definition}[Galois connection]
A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions $f: Y \to X$ and $g: X \to Y$
between posets satisfying $y \leq g(x) \iff f(y) \leq x$.
\end{definition}

\noindent A Galois connection $f \adj g: X \to Y$ is an adjunction between poset categories, regarding
monotone $f$ and $g$ as functors; $f$ is usually referred to as the \emph{upper} (right) adjoint and $g$ as
the \emph{lower} (left) adjoint. Galois connections are closed under component-wise composition and identities
$\id_X \adj \id_X$.
